### 2016-07-07 Projects/DelCo_2012/Report/DelCo_2012.tex
More changes to Ch 3

The fire suppression series included 18 tests in two single-story experimental structures and one two-story structure. Fourteen tests were conducted in the single story structures (7 tests in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied from straight stream to narrow fog. Two of the tests examined suppression tactics for attic fires in residential dwellings, one in each structure. For the attic tests, a smooth bore nozzle was used. The 4 two-story tests used both water and CAF from a straight-stream nozzle.

@@

The fuel load for experiments 11 through 14 consisted of one bale of hay, twelve pallets, eleven and a half sheets plywood on the walls and the wood ``floor assembly". Two stacks of pallets with hay were used as the first items ignited in each experiment. Each stack was composed of six pallets with a half bale of hay. The hay was layered between each pallet. Figure~\ref{fig:Wood_Fuel_Load} shows the pallets prior to ignition. The pallets and hay were weighed prior to each experiment. For the four pallet experiments, the average mass of the two stacks of pallets and the bale of hay was 232.4~kg (511.3~lb).
\begin{figure}[!ht]
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/Wood_Fuel_Package}
	\caption{Photograph of Southeast corner of burn room with wood fuel load.}
	\label{fig:Wood_Fuel_Load}
\end{figure}

@@

\subsection{Attic Fires}

The fuel load for the attic fires, experiments 9 and 10, was essentially the OSB roof underlayment, and the wood trusses that support the roof.  A 1.2 m (4 ft) square hole was opened between the center of the southern room in the single story structure and the attic space. A single stack of pallets as described for the wood fuel package tests was loacted below the hole and ignited.  The fire was allowed to spread into the attic space.  In addition to the connecting hole between the room and the attic space, the attic had openings along the eave line on both sides of the structure and gable openings on each end of the attic space.

@@

In June of 2015, the larger compartment experiments, experiments 15 through 22, were conducted in the lower level of the two story structure.  Two different fuel packages were utitlized, the difference between the furniture fuel package used for experiments 15 and 16, and the furnitiure and pallet fuel package used in experiments 17 through 22 was the addition of two stacks of pallets.

@@

In June 2015 two stacks of pallets were added to the fuel load which consisted of three sofas, OSB lining on the ceiling and walls and the polypropylene carpeting over puf padding.  Each stack had 10 pallets.  The mass of each stack was approximately 175 kg .    When hay was added for ignition, the bales of hay averaged 12 kg.

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB7}
	\caption{Gas Cooling Spray Density Test 7, 30 $^{\circ}$ Fog 
	\label{fig:Burn_Building_Test_7}
\end{figure}

@@

The 88 gas cooling events were completed as part of 9 separate burn sequences in the burn building. For detailed discussion on each of the individual events, please refer to a prior analysis conducted by Mitchell \cite{Mitchell:1}. While Mitchell examined each individual experiment in detail, the results were mixed with CAF showing slightly better or slightly worse than plain water for different gas cooling experimental arrangements.  Hence the conclusions was that there was no significant difference in gas cooling capabilities between CAF and water.

Using the same data, the focus of this analysis was to look at trends within the aggregate data. During each test, the fire was ignited and then developed until the temperature 1.83~m (6~ft) below the ceiling in the gas cooling room reached or exceeded 250~$^{\circ}$C (482~$^{\circ}$F). Once this criteria was met, gas cooling began. Figure~\ref{fig:gas_cooling_exp4} shows 17 cooling events at one thermocouple array during a continuous burn sequence.

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Gas_Cooling/GCSeries4_TC_A3}
	\caption{Time Series of Multiple Gas Cooling Events From 1 of 9 Burn sequences}
	\label{fig:gas_cooling_exp4}
\end{figure}

There were three thermocouple arrays in the gas cooling room (adjacent to the burn room in Fig.~\ref{fig:Gas_Cooling_Instrumentation_Dimensions}). Figures~\ref{fig:gas_cooling_sub1} and \ref{fig:gas_cooling_sub5} show the variation of the impact of the suppression stream on cooling location for three consecutive water and CAF tests, respectively. For both sets a solid 7/8~in nozzle in the middle position was used. The data shows that thermocouples from Gas Cooling Array 2 (GC_2 in the legend), the middle plot of both figures, has more significant temperature decreases than the other two arrays (GC_1 and GC_3); this is the result of Gas Cooling Array 2 being directly ``hit'' by the suppression stream.  Gas Cooling Array 1 was exhibits some amount of suppression agent impact.  Gas Cooling Array 3 is located behind the area where the stream impacts the celing for both the middle and back nozzle positions.  As a result it remains dry and it seems to provide the best representation of the gas cooling.

@@

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for the 30 replicate tests, 15 water and 15 CAF, for all streams aimed at all positions.

@@

The data can be separated to examine the impact of nozzle position (back or middle position) on the ability to cool gases for each of the suppression agents. Figures~\ref{fig:CAFS_Water_mid} and \ref{fig:CAFS_Water_full} show the cooling comparison for 12 replicates in the back position and 3 replicates in the middle position respectively.

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouples in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3).
\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/PDFs/West_Structure/DelCo_2012_West_Structure_Instrumentation2}
	\caption{Schematic of  the Gas Cooling Experiments.}
	\label{fig:Gas_Cooling_Instrumentation_Dimensions2}
\end{figure}
Based on the monitor position (arrow at the doorway in Fig.~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2}), thermocouple arrays A1 and A2 had more direct impact from the suppression agent compared to thermocouple array A3. Figure~\ref{fig:Burn_Building_Test_18b} shows the distribution of the agent within the burn building from a spray density test for a straight stream with CAF at the mid position compared to the positions of the thermocouple trees.
\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Bars/BB18b}
	\caption{Gas Cooling Spray Density Test 18}
	\label{fig:Burn_Building_Test_18b}
\end{figure}

There is a difference in the amount of direct impact from the suppression agent at the three spatial locations. Therefore, the maximum temperature differences for each thermocouple on each of the three arrays should also be analyzed separately. Figures~\ref{fig:CAFS_Water_A1_all}, \ref{fig:CAFS_Water_A2_all}, and \ref{fig:CAFS_Water_A3_all} show the temperature cooling scatter for thermocouple trees A1, A2, and A3 respectively. Note that thermocouple array A3, the array which has the least amount of direct suppression, has the closest fit compared to the solid black line that represents uniform impact for both suppression agents. The fit lines for A1 and A3 also fall within the experimental uncertainty, meaning that differences in the impact of CAF and water are indistinguishable. Thermocouple tree A2, which gets the most significant direct impact from the suppression agents, shows a bias towards water being more effective than CAF.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A1_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A1 Replicates}
	\label{fig:CAFS_Water_A1_all}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A2_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A2 Replicates}
	\label{fig:CAFS_Water_A2_all}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A3_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A3 Replicates}
	\label{fig:CAFS_Water_A3_all}
\end{figure}

@@

The thermocouple array data can also also be decomposed as a function of nozzle position: 12 middle position replicates and 3 back position replicates. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position.

@@

There were 14 suppression tests conducted in the single story purpose-built concrete structure. Four of the tests used a wood pallet fuel load (Section~\ref{sec:fire_suppression_pallet_fuel}) and the remaining tests used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). Ignition occurred in the rear of the single story structures and the available ventilation was through the front doorway and for several tests a rear window was opened to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate room and contents fires typically seen within residential structures. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (room nozzle). Each nozzle flowed 15~s of water or compressed air foam depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling as the stream was not aimed at the seat of the fire (indirect interior attack). The room nozzle was intended to simulate a crew which had advanced into a structure, entered the fire room, positioned just to the side of the entry door, and began suppression on the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

\begin{table}[!ht]
\centering
\caption{Single Story Fire Suppression Test Descriptions}\label{tab:Test_Descriptions}
\begin{tabular}{clllll}
\toprule[1.5pt]
Test $\#$  & Name	& Stream			& Flow Rate		& Fuel           & Ventilation     \\
\midrule
 1 & CAF~1    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door            \\
 2 & Water~1  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door            \\
 3 & CAF~2    &  Solid Stream       &  120~gpm/60~cfm   & Furniture      & Door            \\
 4 & Water~2  &  Solid Stream       &  120~gpm    		& Furniture      & Door            \\
 5 & Water~3  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 6 & CAF~3    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 7 & Water~4  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 8 & CAF~4    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 9 & Attic~1  &  Straight Stream  	&  120~gpm    		& Furniture      & Door            \\
10 & Attic~2  &  Straight Stream  	&  120~gpm/60~cfm   & Furniture      & Door            \\
11 & Water~5  &  30$^{\circ}$ Fog  	&  120~gpm   		& Wood           & Door + Window   \\
12 & CAF~5    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
13 & Water~6  &  30$^{\circ}$ Fog  	&  120~gpm    		& Wood           & Door + Window   \\
14 & CAF~6    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

This study focused on two key types of fire experiments; gas cooling and fire suppression. In each type of experiment a comparison was conducted between plain water and compressed air foam being discharged from nozzles of similar type with similar mass flow rates. In both types of experiments there were no measurable differences between the use of plain water or compressed air foam. These results were not surprising to some of the workshop attendees. However other workshop attendees, that use CAF very successfully in their fire departments, had additional questions. The research team worked to address those follow-up questions by conducting additional fire suppression experiments, one where only wood based fuels were used and another where a larger space with more ventilation was used.  This section will discuss the results and provide potential explanations to support them.

@@

The most widely used of the water-based agents are the foams used on Class B (liquid hydrocarbon) fires. Agents designed primarily for Class A (ordinary combustible) fires have been used most extensively in conjunction with wildland fires. More recently these agents have been promoted for use on a wider range of Class A and in some cases Class B and D (combustible metal) fires. These agents are frequently claimed to be more effective than plain water while being environmentally safe. In some cases, they are also claimed to reduce the quantity and toxicity of smoke. To gain an understanding of how additives might enhance the fire suppression capabilities of water, it is important to examine the principles of fire suppression.

The gas cooling experiments were designed to measure the ability of water streams and CAF streams to cool a hot gas layer with a minimum temperature of 250~$^{\circ}$C (482~$^{\circ}$F) at positions 1.83~m (6~ft) below the ceiling. During the planning of the experiments, some firefighters thought that the excellent heat absorbing capacity of plain water would exceed the heat absorbing capabilities of CAFS, since the injected air, air being a great insulator, would interfere and result in less cooling. It was the belief that the reduced gas cooling capabilities of CAF may have lead to firefighter injuries during interior fire attacks. Other firefighters thought that the bubble structure would provide the same mass flowrate of water with more surface area, which would enable a faster heat transfer rate and therefore, improve the cooling of the hot gas layer. There were also thoughts that the CAFS would coat the ceiling and eliminate energy feedback from the ceiling to the hot gas layer, thereby providing enhanced cooling.

>>>>

The fire suppression series included 20 experiments in the single-story experimental structures and one two-story structure. Tweleve experiments were conducted in the single story structures (6 in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied and included; straight stream, solid stream, and narrow fog.  The two-story tests used both water and CAF from a straight-stream nozzle.

@@

The fuel load for experiments 9 through 12 consisted of one bale of hay, twelve pallets, eleven and a half sheets plywood on the walls and the wood ``floor assembly". Two stacks of pallets with hay were used as the first items ignited in each experiment. Each stack was composed of six pallets with a half bale of hay. The hay was layered between each pallet. Figure~\ref{fig:Wood_Fuel_Load} shows the pallets prior to ignition. The pallets and hay were weighed prior to each experiment. For the four pallet experiments, the average mass of the two stacks of pallets and the bale of hay was 232.4~kg (511.3~lb).
\begin{figure}[!ht]
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/Wood_Fuel_Package}
	\caption{Photograph of Southeast corner of burn room with wood fuel load.}
	\label{fig:Wood_Fuel_Load}
\end{figure}

@@

@@

In June of 2015, the larger compartment experiments, experiments 13 through 20, were conducted in the lower level of the two story structure.  Two different fuel packages were utitlized, the difference between the furniture fuel package used for experiments 13 and 14, and the furnitiure and pallet fuel package used in experiments 15 through 20 was the addition of two stacks of pallets.

@@

In June 2015 two stacks of pallets were added to the fuel load which consisted of three sofas, OSB lining on the ceiling and walls and the polypropylene carpeting over puf padding.  Each stack had 10 pallets.  The mass of each stack was approximately 175 kg.  When hay was added for ignition, the bales of hay averaged 12 kg.

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB7}
	\caption{Gas Cooling Spray Density Test 7, 30 $^{\circ}$ Fog} 
	\label{fig:Burn_Building_Test_7}
\end{figure}

@@

The 88 gas cooling experiments were completed as part of 9 separate burn sequences in the burn building. For detailed discussion on each of the individual events, please refer to a prior analysis conducted by Mitchell \cite{Mitchell:1}. While Mitchell examined each individual experiment in detail, the results were mixed.  In the different gas cooling scenarios CAF showed slightly better or slightly worse than plain water.  Hence the conclusion was that there was no significant difference in gas cooling capabilities between CAF and water.

Using the same data, the focus of this analysis was to look at trends within the aggregate data. During each test, the fire was ignited and then developed until the temperature 1.83~m (6~ft) below the ceiling in the gas cooling room reached or exceeded 250~$^{\circ}$C (482~$^{\circ}$F).  This condition was evidence that a hot gas layer at least 1.83~m deep had filled the room next to the burn room.  Once this criteria was met, gas cooling began. Figure~\ref{fig:gas_cooling_exp4} shows 17 cooling events at one thermocouple array during a continuous burn sequence.  No water or CAF was flowed into the burn room.

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Gas_Cooling/GCSeries4_TC_A3}
	\caption{Time Series of Multiple Gas Cooling Eexperiments from 1 of 9 Burn sequences}
	\label{fig:gas_cooling_exp4}
\end{figure}

There were three thermocouple arrays in the gas cooling room (adjacent to the burn room in Fig.~\ref{fig:Gas_Cooling_Instrumentation_Dimensions}). Figures~\ref{fig:gas_cooling_sub1} and \ref{fig:gas_cooling_sub5} show the variation of the impact of the suppression stream on cooling location for three consecutive water and CAF tests, respectively. For both sets a solid 7/8~in nozzle in the middle position was used. The data shows that thermocouples from Gas Cooling Array 2 (GC\_2 in the legend), the middle plot of both figures, has more significant temperature decreases than the other two arrays (GC\_1 and GC\_3); this is the result of Gas Cooling Array 2 being directly ``hit'' by the suppression stream.  Gas Cooling Array 1 was exhibits some amount of suppression agent impact.  Gas Cooling Array 3 is located behind the area where the stream impacts the celing for both the middle and back nozzle positions.  As a result it remains dry and it seems to provide the best representation of the gas cooling.

@@

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments, 15 water and 15 CAF, which were directly comparable.  These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions.  While the total number of experiments was 88, there where a number of experiments that were performed which did not have a directly comparable experiments between water and CAF.  For example a set of 12 Class A solutions experiments were conducted.  Gas cooling experiments were conducted with water applied on a wide fog nozzle setting, no significant differences were noted so a comparable series with wide fog CAF was not conducted.  Similarly larger water flows were examined but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are not included in this comparison.  The Mitchell report does include all of the data for the 88 experiments \cite{Mitchell:1}

\textbf{Craig - the table of 30 could go here and the R squared paragraph could go with the first discussion of the red line below}

@@

The data can be separated to examine the impact of nozzle position (back or middle position) on the ability to cool gases for each of the suppression agents. Figures~\ref{fig:CAFS_Water_mid} and \ref{fig:CAFS_Water_full} show the cooling comparison for 12 pairs of experiments in the back position and 3 pairs of experiments in the middle position respectively.

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3).  Ideally, the thermocouple is providing an estimate of the temperature of the gas that it is surrounded by.  However, we must remember that the thermocouple is actually providing us with its own temperature.  So if a droplet of water or some CAF cling to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/PDFs/West_Structure/DelCo_2012_West_Structure_Instrumentation2}
	\caption{Schematic of  the Gas Cooling Experiments.}
	\label{fig:Gas_Cooling_Instrumentation_Dimensions2}
\end{figure}
Based on the monitor position (arrow at the doorway in Fig.~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2}), thermocouple arrays A1 and A2 had more direct impact from the suppression agent compared to thermocouple array A3. As an example, Figure~\ref{fig:Burn_Building_Test_18b} shows the distribution of the agent within the burn building from a spray density test for a straight stream with CAF at the mid position compared to the positions of the thermocouple trees.
\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Bars/BB18b}
	\caption{Gas Cooling Spray Density Test 18, CAF, Straight Stream}
	\label{fig:Burn_Building_Test_18b}
\end{figure}

There is a difference in the amount of direct impact from the suppression agent at the three spatial locations. Therefore, the maximum temperature differences for each thermocouple on each of the three arrays should also be analyzed separately. Figures~\ref{fig:CAFS_Water_A1_all}, \ref{fig:CAFS_Water_A2_all}, and \ref{fig:CAFS_Water_A3_all} show the temperature cooling scatter for thermocouple trees A1, A2, and A3 respectively. Note that thermocouple array A3, the array which has the least amount of direct impact from the suppression agents, has the closest fit compared to the solid black line that represents uniform impact for both suppression agents. The fit lines for A1 and A3 also fall within the experimental uncertainty, meaning that differences in the impact of CAF and water are indistinguishable. Thermocouple tree A2, which gets the most significant direct impact from the suppression agents, shows a bias towards water being more effective than CAF on cooling the thermocouples.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A1_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A1 Experimental Pairs}
	\label{fig:CAFS_Water_A1_all}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A2_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A2 Experimental Pairs}
	\label{fig:CAFS_Water_A2_all}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_A3_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Thermocouple Tree A3 Experimental Pairs}
	\label{fig:CAFS_Water_A3_all}
\end{figure}

@@

The thermocouple array data can also also be decomposed as a function of nozzle position: 12 middle position experimental pairs and 3 back experimental pairs. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position.  Again note that the best fit occurs with the data from thermocouple array 3 for both nozzle positions.  This analysis of the data results in the same conclusions at the analysis from the Mitchell paper \cite{Mitchell:1}, the gas cooling performance of CAF, under these conditions, is similar to water.

@@

There were 12 suppression tests conducted in the single story purpose-built concrete structure. Four of the tests used a wood pallet fuel load (Section~\ref{sec:fire_suppression_pallet_fuel}) and the remaining tests used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). Ignition occurred in the rear of the single story structures and the available ventilation was through the front doorway and for several tests a rear window was opened to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate room and contents fires typically seen within residential structures. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (room nozzle). Each nozzle flowed 15~s of water or compressed air foam depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling as the stream was not aimed at the seat of the fire (indirect interior attack). The room nozzle was intended to simulate a crew which had advanced into a structure, entered the fire room, positioned just to the side of the entry door, and began suppression on the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

\begin{table}[!ht]
\centering
\caption{Single Story Fire Suppression Test Descriptions}\label{tab:Test_Descriptions}
\begin{tabular}{clllll}
\toprule[1.5pt]
Test $\#$  & Name	& Stream			& Flow Rate		& Fuel           & Ventilation     \\
\midrule
 1 & CAF~1    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door            \\
 2 & Water~1  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door            \\
 3 & CAF~2    &  Solid Stream       &  120~gpm/60~cfm   & Furniture      & Door            \\
 4 & Water~2  &  Solid Stream       &  120~gpm    		& Furniture      & Door            \\
 5 & Water~3  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 6 & CAF~3    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 7 & Water~4  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 8 & CAF~4    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 9 & Water~5  &  30$^{\circ}$ Fog  	&  120~gpm   		& Wood           & Door + Window   \\
10 & CAF~5    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
11 & Water~6  &  30$^{\circ}$ Fog  	&  120~gpm    		& Wood           & Door + Window   \\
12 & CAF~6    &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

This study focused on two key types of fire experiments; gas cooling and fire suppression. In each type of experiment a comparison was conducted between plain water and compressed air foam being discharged from nozzles of similar type with similar mass flow rates. In both types of experiments there were no significant measurable differences between the use of plain water or compressed air foam. These results were not surprising to some of the technical panel members. However other members, that use CAF very successfully in their fire departments every day, had additional questions. The research team worked to address those follow-up questions by conducting additional fire suppression experiments.  One experiemntal used only wood based fuels, to examine the value of the reduced surface tension of the water in the CAF and its interaction with the surface of the wood fuel.  A final experimental series, used a larger fire room, with more ventilation.  In addition, the nozzle was hand held and could be moved as needed to effectively extinguish the fire.  This section will discuss the results and provide potential explanations to support them.

@@

One of the most widely used of the water-based agents are the foams used on Class B (liquid hydrocarbon) fires. Agents designed primarily for Class A (ordinary combustible) fires have been used most extensively in conjunction with wildland or wildland urban interface fires.  Some of these agents are said to be more effective than plain water while being environmentally safe. In some cases, they are also claimed to reduce the quantity and toxicity of smoke. To gain an understanding of how additives might enhance the fire suppression capabilities of water, it is important to examine the principles of fire suppression.

The gas cooling experiments were designed to measure the ability of water streams and CAF streams to cool a hot gas layer with a minimum temperature of 250~$^{\circ}$C (482~$^{\circ}$F) at positions 1.83~m (6~ft) below the ceiling. During the planning of the experiments, some workshop attendees thought that the excellent heat absorbing capacity of plain water would exceed the heat absorbing capabilities of CAFS, since the injected air, air being a great insulator, would interfere and result in less cooling. It was the belief that the reduced gas cooling capabilities of CAF may have led to firefighter injuries during interior fire attacks. Other firefighters thought that the bubble structure would provide the same mass flowrate of water with more surface area, which would enable a faster heat transfer rate and therefore, improve the cooling of the hot gas layer. There were also thoughts that the CAFS would coat the ceiling and eliminate energy feedback from the ceiling to the hot gas layer, thereby providing enhanced cooling.
